---
title: "Propensity Score"
output: html_notebook
---


```{r Install needed libraries and Packages}
library(ggplot2)
library(reshape2)
library(dbplyr)
library(readr)
library(haven)
library(plyr)
library(knitr)
library(epitools)
library(stargazer)
library(MASS)
library(questionr)
library(lmtest)
library(zoo)
library(MatchIt)
library(Rcpp)
library(gridExtra)
```

```{r Import Data}
#Data is saved as a Stata File
FHS = "https://github.com/copelaa/Framingham-Heart-Study-FHS-Exploration/blob/master/framinghamdata.dta?raw=true"
FHS = read_stata(FHS)

#Take a look at the imported data, and sunderstand the number of rows, columns and column names
head(FHS)
nrow(FHS)
ncol(FHS)
names(FHS)
```

Crude Odds Ratio for the effect of anti-hypertensive medication on incidence of stroke among those with hypertension at the baseline exam?
Crude Odds Ratio
$$
ln(Stroke)=\beta_0+\beta_1(antihypertensive medication)+\varepsilon_i
$$
Note: be sure to exclude individuals who already had a previous stroke at the baseline exam.)
```{r Crude Odd Ratio}
FHS_Prop<-FHS[FHS$prevstrk1 != 1, ]
FHS_Prop<-FHS_Prop[FHS_Prop$prevhyp1 != 0, ]

stroke_lm<- glm(stroke~ bpmeds1, data = FHS_Prop, family= binomial(link = "logit"))

#Odds Ratio
stroke_lm_or<- as.data.frame(exp(cbind(coef(stroke_lm), confint(stroke_lm))))
#Rename Columns + row
names(stroke_lm_or)[1] <- "Odds Ratio"
names(stroke_lm_or)[2] <- "Lower 95% CI"
names(stroke_lm_or)[3] <- "Upper 95% CI"

suppressWarnings(stargazer(stroke_lm, stroke_lm_or,summary=FALSE, type="text", title = "Log Odds and Odds Ratio"))

```
The Odds of Stroke for individuals that were taking anti hypertensive medication at baseline is 1.9 times greater than those who were not, excluding those who had a previous stroke at baseline.

Now we want to control for some factors at baseline as they may affect whether or not an individual uses anti-hypertensive medication as well as their risk of having a stroke. 

```{r Multivariate}

stroke_lm_mv<- glm(stroke~ bpmeds1+age1+sysbp1+diabp1+cursmoke1+totchol1+bmi1+diabetes1, data = FHS_Prop, family= binomial(link = "logit"))

#Odds Ratio
stroke_lm_mv_or<- as.data.frame(exp(cbind(coef(stroke_lm_mv), confint(stroke_lm_mv))))
#Rename Columns + row
names(stroke_lm_mv_or)[1] <- "Odds Ratio"
names(stroke_lm_mv_or)[2] <- "Lower 95% CI"
names(stroke_lm_mv_or)[3] <- "Upper 95% CI"

suppressWarnings(stargazer(stroke_lm_mv, stroke_lm_mv_or,summary=FALSE, type="text", title = "Log Odds and Odds Ratio"))
```

From the above it looks like diabetes1,cusmoke1, diabp1 and age1 may be the most significant factors. 

Propensity score = probability of receiving the treatment (or intervention) as a function of confounders. With the belief that patients with similar propensity scores should show similar distributions of confounders within treatment groups (as if treatment had been randomized) 

```{r Propensity score}

ps_stroke_lm<- glm( bpmeds1~ age1+sysbp1+diabp1+cursmoke1+totchol1+bmi1+diabetes1, data = FHS_Prop, family= binomial(link = "logit"))

#Propensity score

ps_stroke <- data.frame(pr_score = predict(ps_stroke_lm, type = "response"),
                     bpmeds1 =ps_stroke_lm$model$bpmeds1)
head(ps_stroke)
```
After estimating the propensity score, let's plot histograms of the estimated propensity scores by treatment status:

```{r Propensity Score Plot}
labs <- paste( c("antihypertensive medication", "No antihypertensive medication"))
ps_stroke %>%
  mutate(bpmeds1 = ifelse(bpmeds1 == 1, labs[1], labs[2])) %>%
  ggplot(aes(x = pr_score)) +
  geom_histogram(color = "white") +
  facet_wrap(~bpmeds1) +
  xlab("Probability of taking antihypertensive medication") +
  theme_bw()
```

# Executing a matching algorithm
```{r Matching}
#drop NAs
FHS_nomiss<-FHS_Prop[!is.na(FHS_Prop$bpmeds1) & !is.na(FHS_Prop$sysbp1) & !is.na(FHS_Prop$diabp1) & !is.na(FHS_Prop$cursmoke1) & !is.na(FHS_Prop$age1) & !is.na(FHS_Prop$totchol1) & !is.na(FHS_Prop$bmi1) & !is.na(FHS_Prop$diabetes1),]


ps_match<- MatchIt::matchit( bpmeds1~ age1+sysbp1+diabp1+cursmoke1+totchol1+bmi1+diabetes1, data = FHS_nomiss, method= "nearest")

dta_sum<- summary(ps_match)
plot(ps_match)
dta_dist<- dta_sum$sum.all["distance"]

#Create dataframe containing only matched observations 
dta_m<-match.data(ps_match)
dim(dta_m)

```
Note that the final dataset is smaller than the original: it contains 264 observations, meaning that 132 pairs of treated and control observations were matched. Also note that the final dataset contains a variable called distance, which is the propensity score.

#Examining covariate balance in the matched sample 

```{r Visual inspection}
fn_bal <- function(dta, variable) {
  dta$variable <- dta[, variable]
  support <- c(min(dta$variable), max(dta$variable))
  ggplot(dta, aes(x = distance, y = variable, color = bpmeds1)) +
    geom_point(alpha = 0.2, size = 1.3) +
    geom_smooth(method = "loess", se = F) +
    xlab("Propensity score") +
    ylab(variable) +
    theme_bw() +
    ylim(support)
}

grid.arrange(
   fn_bal(FHS_nomiss, "age1"),
   fn_bal(FHS_nomiss, "bmi1") + theme(legend.position = "none"),
   fn_bal(FHS_nomiss, "totchol1"),
   fn_bal(FHS_nomiss, "cursmoke1") + theme(legend.position = "none"),
   fn_bal(FHS_nomiss, "diabp1"),
   fn_bal(FHS_nomiss, "sysbp1") + theme(legend.position = "none"),
   fn_bal(FHS_nomiss, "diabetes1"),
   nrow = 4, widths = c(1, 0.8)
)
   
```